### 2024-07-24 sankey.tex
Mass gain not at 90 degrees

\documentclass[11pt]{article}
\usepackage{sankey}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{ifthen}
\pagestyle{empty}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}

\selectcolormodel{cmyk} % Can be rbg or cmyk, cmyk is better for printing

\begin{tikzpicture}[remember picture, overlay, shift={(5,-5)}]

  \begin{sankeydiagram}
    %% \sankeyset{debug=true}

    \colorlet{water}[rgb]{cyan!75!blue!50!white}
    \colorlet{ice}[rgb]{gray!20!gray!20!gray!20}
    %% \colorlet{air}[rgb]{gray!1!gray!1!gray!1}
    \colorlet{air}{yellow!50}
    \colorlet{red}[rgb]{red!50!orange!75!white}
    \colorlet{black}[rgb]{black!100!black!100!black!100}
    \colorlet{gray}[rgb]{black!30!black!30!black!30}

    \sankeyset{
      %% comment for gray outline
      draw/.style={
        draw=white,
        line width=0pt,line cap=round,line join=round,
      },
      start style=arrow,end style=arrow,
      %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
      ratio=1mm/60,
      minimum radius=0mm,
      water/.style={fill/.style={draw=water,line width=0pt,fill=water,},},
      ice/.style={fill/.style={draw=ice,line width=0pt,fill=ice,},},
      air/.style={fill/.style={draw=air,line width=0pt,fill=air,},},
      black/.style={fill/.style={draw=black,line width=0pt,fill=black,},},
      red/.style={fill/.style={fill=none, top color=red, bottom color=red},},
      ice-to-water/.style={fill/.style={fill=none, left color=ice, right color=water},},
      ice-to-air/.style={fill/.style={fill=none, left color=ice, right color=air},},
      ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
      ice-to-air-vert/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
      ice-to-water-vert/.style={fill/.style={fill=none, bottom color=ice, top color=water},},
      water-to-air-vert/.style={fill/.style={fill=none, right color=water, left color=air},},
      water-to-air/.style={fill/.style={fill=none, left color=water, right color=air},},
      air-to-ice/.style={fill/.style={fill=none, left color=air, right color=ice},},
      red-to-ice/.style={fill/.style={fill=none, left color=red, right color=ice},},
      water-to-ice/.style={fill/.style={fill=none, left color=water, right color=ice},},
      black-to-ice/.style={fill/.style={fill=none, bottom color=ice, top color=black},},
      in-black/.style={draw/.style={draw=gray,thick=3},},
      out-red/.style={draw/.style={draw=red,thick=3},},
      debug label/.style={
        overlay,
        draw,
        font=\tiny,
        text=debug color,text opacity=1,
        inner sep=.1em,
        fill=white,fill opacity=0.5,
        rounded corners=.1em,
        %% node contents={\name (\qty)},
        node contents={\name},
      },
    }

    % storage of labels
    \newcommand\LabSet[2]{% node name, label
      \expandafter\edef\csname#1@Lab\endcsname{#2}}
    \newcommand\Lab[1]{% node name
      \csname#1@Lab\endcsname}
    % storage of quantities
    \newcommand\QSet[2]{% node name, quantity
      \expandafter\edef\csname#1@Q\endcsname{\fpeval{#2}}}
    \newcommand\Q[1]{% node name
      \csname#1@Q\endcsname}
    % all nodes with their name, label and quantity
    \newcommand\qq[1]{\sankeygetnodeqty{#1}}

    \newcommand\labelleft[1]{ % valname
      \node[font=\tiny, anchor=east]
      at ([yshift=0mm,xshift=-3mm]#1.center) {\Lab{v#1}: \qq{#1}};
    }
    \newcommand\labelright[2]{ % valname
      \node[font=\tiny, anchor=west]
      at ([yshift=0mm,xshift=3mm]#1.center) {\qq{#1}: \Lab{#2}};
    }

    \sankeyset{
      def data/.code args={#1/#2/#3}{% node name/label/values
        \LabSet{#1}{#2}
        \QSet{#1}{#3}
        \typeout{#1: \Q{#1} Gt (\Lab{#1})}
      },
      def data/.list={
        %% INSERT_CSV_HERE
        %% COMPUTED
        {vML/Mass loss/
          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})-
          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})},
        {vMG/Mass gain/
          (\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}+\Q{vFRGAIN})-
          (\Q{vSU}+\Q{vRU}+\Q{vBM}+\Q{vDYN}+\Q{vFRLOSS}+\Q{vGZRET}+\Q{vEV})}
      },
    }

    \sankeyset{ice} % default color until otherwise specified
    \newcommand\yoff{2mm}

    \sankeynode{name=SF, quantity=\Q{vSF}}
    \sankeynode{name=DP, quantity=\Q{vDP}, at={[yshift=\yoff]SF.north}, anchor=right}
    \sankeynode{name=CD, quantity=\Q{vCD}, at={[yshift=\yoff]DP.north}, anchor=right}
    \sankeynode{name=RF, quantity=\Q{vRF}, at={[yshift=\yoff]CD.north}, anchor=right}

    \labelleft{SF}
    \labelleft{DP}
    \labelleft{CD}
    \labelleft{RF}

    \sankeystart{SF}
    \sankeystart[air]{DP}
    \sankeystart{CD}
    \sankeystart[water]{RF}

    \sankeynode{name=f0, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
      at={[xshift=15mm, yshift=0mm]SF.right}, anchor=right,
      forked={\Q{vRF}/RFi, \Q{vCD}/CDi, \Q{vDP}/DPi, \Q{vSF}/SFi}}

    \sankeyoutin{SF}{SFi}
    \sankeydubins[minimum radius=3mm,air-to-ice]{DP}{DPi}
    \sankeydubins[minimum radius=3mm]{CD}{CDi}
    \sankeydubins[minimum radius=3mm,water-to-ice]{RF}{RFi}

    \ifthenelse{\Q{vMG} > \Q{vML}} {
        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vFRGAIN},
          at={[xshift=15mm]f0.right}, anchor=right,
          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \qq{f0}/f0i}}
    } {
        \sankeynode{name=f1, quantity=\Q{vRFZ}+\qq{f0}+\Q{vML}+\Q{vFRGAIN},
          at={[xshift=15mm]f0.right}, anchor=right,
          forked={\Q{vRFZ}/RFZi, \Q{vFRGAIN}/FG, \Q{vML}/MLi, \qq{f0}/f0i}}
    }

    \sankeyoutin{f0}{f0i}
    \sankeyadvance{f1}{15mm}

    \ifthenelse{\Q{vMG} > \Q{vML}}
    {
      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/f8, \Q{vDYN}/f6,
        \Q{vGZRET}/GZ, \Q{vFRLOSS}/FL, \Q{vBM}/BM, \Q{vMG}/MG}
    }{
      \sankeyfork{f1}{
        %% \Q{vMG}/MG,
        \Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/f8, \Q{vDYN}/f6,
        \Q{vGZRET}/GZ, \Q{vFRLOSS}/FL, \Q{vBM}/BM}
    }

    \sankeyturnleft{RFZo}{1}
    \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}

    \sankeyadvance{f6}{10mm}
    \sankeyfork{f6}{\Q{vICE}/CALV, \Q{vSUB}/SUB}
    \sankeyadvance[ice-to-water]{SUB}{10mm}

    \sankeynode{name=CALVa, quantity=\qq{CALV}, at={[yshift=\yoff]SUB.left}, anchor=right}
    \sankeyoutin{CALV}{CALVa}

    \sankeynode{name=f8a, quantity=\qq{f8}, at={[yshift=\yoff]CALVa.left}, anchor=right}
    \sankeyoutin[ice-to-water]{f8}{f8a}

    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]SUB.right}, anchor=left}
    \sankeyoutin{GZ}{GZa}

    \sankeynode{name=FLa, quantity=\qq{FL}, at={[yshift=-\yoff]GZa.right}, anchor=left}
    \sankeyoutin{FL}{FLa}

    \sankeynode{name=BMa, quantity=\qq{BM}, at={[yshift=-\yoff]FLa.right}, anchor=left}
    \sankeyoutin[ice-to-water]{BM}{BMa}

    \sankeyadvance[out-red]{FLa}{10mm}
    \sankeyend[out-red]{FLa}
    \labelright{FLa}{vFRLOSS}

    \ifnum \numexpr\Q{vMG} > \Q{vML}
      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-35,
        at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
      \node[font=\tiny, anchor=east]
      at ([yshift=0mm,xshift=0mm]MGa.right) {Mass gain: \qq{MGa}};
      \sankeyoutin[ice-to-black]{MG}{MGa}
    \fi

    \sankeyfork{f8a}{\Q{vEV}/EV, \Q{vRU}/RU}
    \sankeynode{name=EVa, quantity=\qq{EV},
      at={[yshift=2mm,xshift=5mm]f8a.left}, anchor=left, angle=45}
    \sankeyoutin[water-to-air]{EV}{EVa}
    \sankeyadvance[air]{EVa}{1mm}
    \sankeyend[air]{EVa}

    \sankeynode{name=SUa, quantity=\qq{SU},
      at={[xshift=-6mm]EVa.left}, anchor=right, angle=45}
    %% \sankeyoutin[ice-to-air{SU}{SUa}
    \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}
    \sankeyend[air]{SUa}

    \ifthenelse{\Q{vMG} < \Q{vML}} { % mass loss
      \sankeynode{name=ML, quantity=\Q{vML},
        at={[yshift=6mm]RF.north}, anchor=right, angle=-45}
      \node[font=\tiny, anchor=east]
      at ([xshift=-3mm,yshift=2mm]ML.west) {Mass loss: \qq{ML}};

      % frontal advance next to drawdown (mass loss)
      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
        at={[yshift=0mm,xshift=6mm]ML.left}, anchor=right, angle=-45}

      \sankeystart[red]{ML}
      \sankeyadvance[red]{ML}{5mm}
      \sankeyoutin[red-to-ice]{ML}{MLi}
    }{ % mass gain
      \sankeyadvance[black]{MGa}{3mm}
      \sankeyend[black]{MGa}
      \sankeynode{name=FGi, quantity=\Q{vFRGAIN},
        at={[xshift=-15mm,yshift=8mm]RFZi.right}, anchor=left, angle=-45}
    }

    \ifnum \numexpr\qq{FGi} > 0
    \sankeystart[in-black]{FGi}
    \sankeyoutin[in-black]{FGi}{FG}
    \node[font=\tiny, anchor=east]
    at ([xshift=-3mm,yshift=3mm]FGi.west) {Frontal advance: \qq{FGi}};
    \fi

    \sankeyadvance[ice-to-water,out-red]{GZa}{10mm}
    \sankeyend[water,out-red]{GZa}
    \sankeyadvance{CALVa}{10mm}
    \sankeyend{CALVa}

    \sankeyset{water}
    \sankeyadvance{RU}{10mm}; \sankeyend{RU}
    \sankeyadvance{BMa}{10mm}; \sankeyend{BMa}
    \sankeyadvance{SUB}{10mm}; \sankeyend{SUB}

    \labelright{RU}{vRU}
    \labelright{BMa}{vBM}
    \labelright{SUB}{vSUB}
    \labelright{CALVa}{vICE}
    \labelright{GZa}{vGZRET}

    \node[font=\tiny, anchor=west]
    at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};

    \node[font=\tiny, anchor=west]
    at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};

    \node[font=\tiny, anchor=east]
    at ([yshift=-0.5mm,xshift=1mm]RFZo.center) {\Lab{vRFZ}: \qq{RFZi}};

  \end{sankeydiagram}
\end{tikzpicture}
\end{document}