### 2025-03-10 sankey.tex
co-author edits and comments incorporated

\documentclass[11pt]{article}
\usepackage{sankey}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{ifthen}
\pagestyle{empty}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}

\selectcolormodel{cmyk} % Can be rbg or cmyk, cmyk is better for printing

\begin{tikzpicture}[remember picture, overlay, shift={(5,-5)}]

  \begin{sankeydiagram}
    %% \sankeyset{debug=true}

    \colorlet{water}[rgb]{cyan!75!blue!50!white}
    \colorlet{ice}[rgb]{gray!20!gray!20!gray!20}
    %% \colorlet{air}[rgb]{gray!1!gray!1!gray!1}
    \colorlet{air}{yellow!50}
    \colorlet{red}[rgb]{red!50!orange!75!white}
    \colorlet{black}[rgb]{black!100!black!100!black!100}
    \colorlet{gray}[rgb]{black!30!black!30!black!30}

    \sankeyset{
      %% comment for gray outline
      %% draw/.style={
      %%   draw=white,
      %%   line width=0pt,line cap=round,line join=round,
      %% },
      %% comment for gray outline
      draw/.style={
        draw=black,
        line width=0.1mm,line cap=round,line join=round,
      },
      start style=arrow,end style=arrow,
      %% fill/.style={fill=yellow!10,line width=0pt,draw=yellow!10},
      ratio=1mm/60,
      minimum radius=0mm,
      water/.style={fill/.style={draw=water,line width=0pt,fill=water,},},
      ice/.style={fill/.style={draw=ice,line width=0pt,fill=ice,},},
      air/.style={fill/.style={draw=air,line width=0pt,fill=air,},},
      black/.style={fill/.style={draw=black,line width=0pt,fill=black,},},
      red/.style={fill/.style={fill=none, top color=red, bottom color=red},},
      ice-to-water/.style={fill/.style={fill=none, left color=ice, right color=water},},
      ice-to-air/.style={fill/.style={fill=none, left color=ice, right color=air},},
      ice-to-black/.style={fill/.style={fill=none, left color=ice, right color=black},},
      ice-to-air-vert/.style={fill/.style={fill=none, top color=ice, bottom color=air},},
      ice-to-water-vert/.style={fill/.style={fill=none, bottom color=ice, top color=water},},
      water-to-air-vert/.style={fill/.style={fill=none, right color=water, left color=air},},
      water-to-air/.style={fill/.style={fill=none, left color=water, right color=air},},
      air-to-ice/.style={fill/.style={fill=none, left color=air, right color=ice},},
      red-to-ice/.style={fill/.style={fill=none, left color=red, right color=ice},},
      water-to-ice/.style={fill/.style={fill=none, left color=water, right color=ice},},
      black-to-ice/.style={fill/.style={fill=none, bottom color=ice, top color=black},},
      in-black/.style={draw/.style={draw=gray,thick=3},},
      out-red/.style={draw/.style={draw=red,thick=3},},
      out-blue/.style={draw/.style={draw=water,thick=3},},
      debug label/.style={
        overlay,
        draw,
        font=\tiny,
        text=debug color,text opacity=1,
        inner sep=.1em,
        fill=white,fill opacity=0.5,
        rounded corners=.1em,
        %% node contents={\name: \qty},
        node contents={\qty},
        %% node contents={\name},
      },
    }

    % storage of labels
    \newcommand\LabSet[2]{% node name, label
      \expandafter\edef\csname#1@Lab\endcsname{#2}}
    \newcommand\Lab[1]{% node name
      \csname#1@Lab\endcsname}
    % storage of quantities
    \newcommand\QSet[2]{% node name, quantity
      \expandafter\edef\csname#1@Q\endcsname{\fpeval{#2}}}
    \newcommand\Q[1]{% node name
      \csname#1@Q\endcsname}
    % all nodes with their name, label and quantity
    \newcommand\qq[1]{\sankeygetnodeqty{#1}}

    \newcommand\labelleft[1]{ % valname
      \node[font=\tiny, anchor=east]
      at ([yshift=0mm,xshift=-3mm]#1.center) {\Lab{v#1}: \qq{#1}};
    }
    \newcommand\labelright[2]{ % valname
      \node[font=\tiny, anchor=west]
      at ([yshift=0mm,xshift=3mm]#1.center) {\qq{#1}: \Lab{#2}};
    }
    \newcommand\labelbold[2]{ % valname
      \node[font=\tiny, anchor=west]
      at ([yshift=0mm,xshift=3mm]#1.center) {\textbf{\qq{#1}: \Lab{#2}}};
    }

    \sankeyset{
      def data/.code args={#1/#2/#3}{% node name/label/values
        \LabSet{#1}{#2}
        \QSet{#1}{#3}
        \typeout{#1: \Q{#1} Gt (\Lab{#1})}
      },
      def data/.list={
        %% INSERT_CSV_HERE
        %% COMPUTED
        {vSMBin/SMB Inputs/\Q{vSF}+\Q{vRF}+\Q{vCD}+\Q{vDP}},
        {vINPUT/Inputs/\Q{vSMBin}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}},
        {vSMBout/SMB Outputs/\Q{vSU}+\Q{vRU}+\Q{vEV}},
        {vDYN/Dynamics/\Q{vCALV}+\Q{vFRONTMELT}+\Q{vSHELFMELT}+\Q{vFRONTRET}+\Q{vGZRET}+\Q{vBMB}},
        {vOUTPUT/Outputs/\Q{vSMBout}+\Q{vDYN}},
        {vSMB/SMB/\Q{vSMBin}-\Q{vSMBout}},
        {vML/Drawdown; Retreat/\Q{vOUTPUT}-\Q{vINPUT}},
        {vMG/Build-up; Expansion/\Q{vINPUT}-\Q{vOUTPUT}}
      },
    }

    \sankeyset{ice} % default color until otherwise specified
    \newcommand\yoff{3mm}

    \sankeynode{name=SF, quantity=\Q{vSF}}
    \sankeynode{name=DP, quantity=\Q{vDP}, at={[yshift=\yoff]SF.north}, anchor=right}
    \sankeynode{name=CD, quantity=\Q{vCD}, at={[yshift=\yoff]DP.north}, anchor=right}
    \sankeynode{name=RF, quantity=\Q{vRF}, at={[yshift=\yoff]CD.north}, anchor=right}

    \sankeystart{SF}; \labelleft{SF}
    \sankeystart[air]{DP}; \labelleft{DP}
    \sankeystart{CD}; \labelleft{CD}
    \sankeystart[water]{RF}; \labelleft{RF}

    \sankeynode{name=SMBIN, quantity=\Q{vSF}+\Q{vDP}+\Q{vCD}+\Q{vRF},
      at={[xshift=15mm, yshift=0mm]SF.right}, anchor=right,
      forked={\Q{vRF}/RFi, \Q{vCD}/CDi, \Q{vDP}/DPi, \Q{vSF}/SFi}}

    \sankeyoutin{SF}{SFi}
    \sankeydubins[minimum radius=3mm,air-to-ice]{DP}{DPi}
    \sankeydubins[minimum radius=3mm]{CD}{CDi}
    \sankeydubins[minimum radius=3mm,water-to-ice]{RF}{RFi}

    \sankeynode{name=SHELFFREEZE, quantity=\Q{vSHELFFREEZE},
      at={[yshift=-\yoff]SF.south}, anchor=left}

    \ifthenelse{\Q{vML} > \Q{vMG}} { % mass loss
      \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
        quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vML}+\Q{vRFZ},
        forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \Q{vML}/MLi,
          \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
    } { % mass gain edge case
      \sankeynode{name=f1, at={[xshift=30mm, yshift=\yoff]SHELFFREEZE.south}, anchor=right,
        quantity=\qq{SMBIN}+\Q{vFRONTADV}+\Q{vSHELFFREEZE}+\Q{vRFZ},
        forked={\Q{vRFZ}/RFZi, \Q{vFRONTADV}/FG, \qq{SMBIN}/SMBINi, \Q{vSHELFFREEZE}/SHELFFREEZEi}}
    }

    \sankeyoutin{SMBIN}{SMBINi}
    \sankeyadvance{f1}{15mm}

    \ifnum \numexpr\Q{vSHELFFREEZE} > 0
    \labelleft{SHELFFREEZE}
    \sankeystart[water]{SHELFFREEZE}
    \sankeydubins[water-to-ice]{SHELFFREEZE}{SHELFFREEZEi}
    \fi

    \ifthenelse{\Q{vML} > \Q{vMG}} { % mass loss
      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/RUEV,
        \Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT, \Q{vSHELFMELT}/SHELFMELT,
        \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB}
    }{ % mass gain edge case (AQ_east)
      \sankeyfork{f1}{\Q{vRFZ}/RFZo, \Q{vSU}/SU, \Q{vRU}+\Q{vEV}/RUEV,
        \Q{vCALV}/CALV, \Q{vFRONTMELT}/FRONTMELT, \Q{vSHELFMELT}/SHELFMELT,
        \Q{vFRONTRET}/FRONTRET, \Q{vGZRET}/GZ, \Q{vBMB}/BMB, \Q{vMG}/MG}
    }
    \sankeyturnleft{RFZo}{1}
    \sankeydubins[minimum radius=5mm,ice-to-water-vert]{RFZo}{RFZi}
    \node[font=\tiny, anchor=east]
           at ([yshift=+1.5mm,xshift=1mm]RFZo.north) {\Lab{vRFZ}: \qq{RFZi}};

    \ifthenelse{\Q{vFRONTMELT} > \Q{vSHELFMELT}} { % Greenland
      %% Discharge splits to FRONTMELT (straight) & CALV (above).
      %% SHELFMELT below FRONTMELT. FRONTRET below SHELFMELT.
      \sankeyadvance{FRONTMELT}{20mm}

      \sankeynode{name=SHELFMELTa,quantity=\qq{SHELFMELT},
        at={[yshift=-\yoff, xshift=10mm]FRONTMELT.right}}
      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
        at={[yshift=-\yoff]SHELFMELTa.right}}
      \sankeyadvance[ice-to-water]{FRONTMELT}{10mm}

      \sankeynode{name=CALVa, quantity=\qq{CALV},
        at={[yshift=\yoff,xshift=0mm]FRONTMELT.left}, anchor=right}

      \sankeyset{water}
      \sankeyadvance{FRONTMELT}{10mm}; \sankeyend{FRONTMELT}
      \labelright{FRONTMELT}{vFRONTMELT}
    } {
      % Antarctica
      %% Discharge splits to SHELFMELT (straight) & CALV (above).
      %% No FRONTMELT. Then FRONTRET
      \sankeyadvance{SHELFMELT}{10mm}
      \sankeynode{name=SHELFMELTa, quantity=\qq{SHELFMELT},
        at={[yshift=0, xshift=20mm]SHELFMELT.right}, anchor=right}

      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
        at={[yshift=-\yoff]SHELFMELT.right}}
      \sankeynode{name=CALVa, quantity=\qq{CALV},
        at={[yshift=\yoff,xshift=0mm]SHELFMELTa.left}, anchor=right}
      \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET},
        at={[yshift=-\yoff]SHELFMELTa.right}, anchor=left}
    }

    \sankeyset{ice}

    \sankeyoutin[ice-to-water]{SHELFMELT}{SHELFMELTa}
    \sankeyadvance[water]{SHELFMELTa}{10mm}; \sankeyend[water]{SHELFMELTa}
    \labelright{SHELFMELTa}{vSHELFMELT}

    %% \sankeynode{name=FRONTRETa, quantity=\qq{FRONTRET}, at={[yshift=-\yoff]SHELFMELTa.right}}
    \sankeyoutin[out-blue]{FRONTRET}{FRONTRETa};

    \sankeynode{name=GZa, quantity=\qq{GZ}, at={[yshift=-\yoff]FRONTRETa.right}, anchor=left}

    \sankeynode{name=BMBa, quantity=\qq{BMB}, at={[yshift=-\yoff]GZa.right}, anchor=left}
    \sankeyoutin[ice-to-water]{BMB}{BMBa}; \sankeyadvance[water]{BMBa}{10mm};
    \sankeyend[water]{BMBa}; \labelright{BMBa}{vBMB}

    \sankeyoutin[ice-to-water]{GZ}{GZa}; \sankeyadvance[water]{GZa}{10mm};
    \sankeyend[water]{GZa}; \labelbold{GZa}{vGZRET}
    \sankeyadvance[out-blue]{FRONTRETa}{10mm}; 
    \sankeyend[out-blue]{FRONTRETa}; \labelbold{FRONTRETa}{vFRONTRET}

    \sankeyadvance{CALV}{5mm};
    \sankeyoutin{CALV}{CALVa};

    %% SMB outputs
    \sankeynode{name=RUEVa, quantity=\qq{RUEV}, at={[yshift=\yoff]CALVa.left}, anchor=right}
    \sankeyoutin[ice-to-water]{RUEV}{RUEVa}
    \sankeyfork{RUEVa}{\Q{vEV}/EV, \Q{vRU}/RU}
    \sankeynode{name=EVa, quantity=\qq{EV},
      at={[yshift=\yoff,xshift=5mm]RUEVa.left}, anchor=left, angle=45}
    \sankeyoutin[water-to-air]{EV}{EVa}; \sankeyadvance[air]{EVa}{1mm}
    \sankeyend[air]{EVa};
    \node[font=\tiny, anchor=west] at ([yshift=0mm,xshift=\yoff]EVa.west) {\qq{EVa}: \Lab{vEV}};

    \sankeyadvance[water]{RU}{10mm}; \sankeyend[water]{RU}; \labelright{RU}{vRU}

    \sankeynode{name=SUa, quantity=\qq{SU}, at={[xshift=-6mm]EVa.left}, anchor=right, angle=45}
    \sankeydubins[minimum radius=6mm,ice-to-air]{SU}{SUa}; \sankeyend[air]{SUa}
    \node[font=\tiny, anchor=west] at ([yshift=4mm,xshift=\yoff]SUa.west) {\qq{SUa}: \Lab{vSU}};

    \sankeyadvance{CALVa}{10mm}; \sankeyend{CALVa}; \labelright{CALVa}{vCALV}

    \ifthenelse{\Q{vMG} < \Q{vML}} { % mass loss
      \sankeynode{name=ML, quantity=\Q{vML},
        at={[yshift=6mm]RF.north}, anchor=right, angle=-45}
      \node[font=\tiny, anchor=east]
      at ([xshift=-3mm,yshift=2mm]ML.west) {\textbf{Drawdown or retreat: \qq{ML}}};

      % frontal advance next to drawdown (mass loss)
      \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
        at={[yshift=0mm,xshift=6mm]ML.left}, anchor=right, angle=-45}

      \sankeystart[red]{ML}
      \sankeyadvance[red]{ML}{5mm}
      \sankeyoutin[red-to-ice]{ML}{MLi}
    }{ % mass gain
      \sankeynode{name=MGa, quantity=\qq{MG}, angle=-35,
        at={[xshift=5mm,yshift=-3mm]MG.right}, anchor=right}
      \node[font=\tiny, anchor=east]
        at ([yshift=0mm,xshift=0mm]MGa.right) {Build-up or expansion: \qq{MGa}};
      \sankeyoutin[ice-to-black]{MG}{MGa}; \sankeyadvance[black]{MGa}{3mm}
      \sankeyend[black]{MGa}
      \sankeynode{name=FRONTADVi, quantity=\Q{vFRONTADV},
        at={[xshift=-15mm,yshift=8mm]RFZi.right}, anchor=left, angle=-45}
    }

    \ifnum \numexpr\qq{FRONTADVi} > 0
      \sankeystart[in-black]{FRONTADVi}
      \sankeyoutin[in-black]{FRONTADVi}{FG}
      \node[font=\tiny, anchor=east]
          at ([xshift=-3mm,yshift=3mm]FRONTADVi.west) {Frontal advance: \qq{FRONTADVi}};
    \fi

    %% % fancy labels
    %% \sankeynode[debug]{name=Discharge, at={[xshift=-10mm]DISCH.north}, anchor=left,
    %%   quantity=\Q{vDISCH}}
    %% \sankeynode[debug]{name=SMB Inputs, at={[xshift=3mm]SMBIN.north}, anchor=left,
    %%   quantity=\Q{vSMBin}}
    %% \sankeynode[debug]{name=SMB Outputs, at={[xshift=0mm]RFZo.south}, anchor=left,
    %%   quantity=\Q{vSMBout}}

  \end{sankeydiagram}
\end{tikzpicture}
\end{document}